---
title: "Introduction to OmniAgeRData"
author:
- name: Zhaozhen Du
  affiliation: Shanghai Institute of Nutrition and Health, CAS
  email: duzhaozhen2022@sinh.ac.cn
- name: Andrew E. Teschendorff
  affiliation: Shanghai Institute of Nutrition and Health, CAS
  email: andrew@sinh.ac.cn
package: OmniAgeRData
output: 
  BiocStyle::html_document:
    toc_float: true
vignette: |
  %\VignetteIndexEntry{Introduction to OmniAgeRData}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This package provides aging omic biomarker model parameters and example data for
the `OmniAgeR` package. It utilizes `ExperimentHub` to deliver both traditional 
clock coefficients and large-scale synthetic datasets.

# Exploring Available Resources
You can explore all available OmniAgeR datasets and models 
through ExperimentHub:

```{r explore, eval=TRUE}
library(OmniAgeRData)
library(ExperimentHub)

eh <- ExperimentHub()
query(eh, "OmniAgeRData")
```

# Data Retrieval
The recommended way to load data is using our wrapper function getOmniAgeData().

```{r retrieve, eval=TRUE}
# Simulate local tests to pass BiocCheck
local_dir <- tempdir()
saveRDS(matrix(1:10), file.path(local_dir, "LungInv.rds"))

lung_data <- getOmniAgeData("LungInv", localTest = TRUE, localDir = local_dir)
head(lung_data)
```
**Note on Large Datasets:** For extremely large datasets like PCClocks_data.qs2 
(2GB), we highly recommend a stable and fast internet connection. If the 
default R download mechanism fails or is too slow, please refer to the 
troubleshooting section below.

# Troubleshooting: Manual Caching for Large Files

If you experience network issues when downloading large .qs2 files, 
you can bypass the R downloader by manually caching the files.

Step 1: Download the dataset directly from our Zenodo repository via your web 
browser or a command-line tool like wget 
(which supports resuming broken downloads).

Step 2: Use the following R code to locate the expected local cache path and 
move your downloaded file there:

```{r manual_cache, eval=FALSE}
# 1. Locate the correct destination path in your local ExperimentHub cache
target_path <- AnnotationHub::cache(query(eh, "PCClocks_data.qs2")[1])

# 2. Copy your manually downloaded file to the cache location
# Replace "path/to/your/downloaded/file" with your actual file path
file.copy("path/to/your/downloaded/file/PCClocks_data.qs2", target_path,
    overwrite = TRUE
)

# 3. Now you can load the data normally in R!
pc_clocks <- getOmniAgeData("PCClocks_data.qs2")
```

# Session Information
```{r sessionInfo, eval=TRUE}
sessionInfo()
```

